---
title: "STAT796_HW1_code"
author: "Colleen Brents"
date: "January 23, 2019"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE
)
```

```{r load libraries, message=FALSE, warning=FALSE}
library(ggplot2)
library(dplyr)
library(tidyr)
library(readr)
library(knitr)
```
##1. Import Data  
```{r import data, message=FALSE, warning=FALSE}

url <- paste0("https://raw.githubusercontent.com/cbrents/stat796logreg/master/hw1/ICU_raw.txt")
icu_raw<-read_table2(url)
```
##2. rename and recode 
```{r eval=TRUE}
icu<- icu_raw %>% 
    transmute(id=ID,
              died=STA,
              age=AGE,
              female=GENDER,
              race=factor(RACE, 
                          levels=1:3,
                          labels=c("white", "black", "other")),
              service=factor(SER,
                             levels=0:1,
                             labels=c("medical", "surgical")),
              cancer=CAN,
              renal=CRN,
              infect=INF,
              cpr=CPR,
              sbp=SYS,
              hr=HRA,
              prv_admit=PRE,
              type=factor(TYP,
                          levels=0:1,
                          labels=c("elective", "emergency")),
              fracture=FRA,
              po2_g60=PO2,
              '1-PO2'= po2_g60,
              pco2_g45=PCO,
              bicarb_g18=BIC,
              '1 - BIC' = bicarb_g18,
              creat_g2=CRE,
              conscious=factor(LOC,
                levels=0:2,
                labels=c("no_coma_or_stupor", "stupor", "coma")))
```
##3. summary stats for entire cohort
###3a. average, max, min, sc of age (years)

```{r eval = TRUE}
x<- icu %>% 
  summarize(n=n(),
            mean_age = mean(age),
            max_age = max(age),
            min_age = min(age),
            sd_age =sd(age)) 
print(x)
```

###3b. count of subjects with each service type 
```{r eval=TRUE}
a<- table (icu$service)
print(a)
``` 
###3c. count of subjects with procedure type  
```{r}
b<- table(icu$type)
print(b)
```

###3d. count of subjects of each race and each sex
```{r}
#count(icu, race, female)
z<- table(icu$race, icu$female)
colnames(z)<- c("male", "female")
print(z)
```

###3e. count of subject by consciousness 
```{r}
#count(icu, conscious)
x<- table(icu$conscious)
print(x)
```

##4. summary stats for cohort, grouped by by vital status
###4a. average, max, min, sd age
```{r eval=TRUE}
age_table <- icu %>%
  group_by(Died=died) %>%
  summarize(n=n(),
           mean_age = mean(age),
            max_age = max(age),
            min_age = min(age),
            sd_age =sd(age))  # died == 0  alive == 1
print(age_table)
```
###4b. count by service type 
```{r}
vitals <- icu %>% 
  group_by(died)
y<- table(vitals$type)
print(y)
```
###4c. count by procedure type 
```{r}
a<- table(vitals$type)
print(a)
```


###4d. count by race and sex 
```{r }
z<- table(vitals$race, vitals$female)
 colnames(z)<- c("male", "female")
print(z)
```
###4e. count by consciousness 
```{r}
x<- table(vitals$conscious)
print(x)
```

##5. graphical summaries
###5a. scatterplot of sbp against age 
```{r}
ggplot(icu) +
  theme_bw() +
  geom_point(aes(x=sbp, 
                 y=age, 
                 color= as.factor(died), 
                 shape=as.factor(died))) +
  xlab("systolic blood pressure") +
  ylab("age (years)") +
  ggtitle("5a - systolic blood pressure vs age") +
  scale_color_discrete(name="Died or not") +
  scale_shape_discrete(name="Died or not")  #having the same text for legens combines them 
```

###5b. boxplots of hr, grouped by consciousness
```{r}
ggplot(icu) +
  theme_bw() +
  geom_boxplot(aes(group = as.factor(died), x=as.factor(died), y= hr)) +
  xlab("vital status at discharge") +
  ylab("heart rate (bpm)") +
  ggtitle("5b - boxplot of hr by consciousness") +
  scale_x_discrete(labels = c("alive", "dead"))
```

###5c. histograms of age 
```{r, message=FALSE, warning=FALSE}
ggplot(icu) +
  theme_bw() +
  geom_histogram(aes(x=age)) +
  xlab("age(years)") +
  ylab("count") +
  ggtitle("5c - histogram of age") 

```

###5d. densit estimates of hr, grouped by consciousness 
```{r graphical summaries of data}
ggplot(icu) + 
  theme_bw() + 
  geom_line(aes(x=hr, col=as.factor(conscious), group=conscious), stat="density") +
  xlab("Age (in years)") +
  ylab("Density") +
  scale_color_discrete("conscious") +
  ggtitle("5d - density estimate of hr, grouped by consciousness") #from notes 
```
