---
title: "hw2"
author: "Colleen Brents"
date: "January 31, 2019"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
options(scipen=1, digits=2)
```
#question about 2e and 3d

```{r load libraries, include=FALSE}
library(ggplot2)
library(dplyr)
library(tidyr)
library(readr)
library(knitr)
#library(kableExtra)
```

```{r data input, include=FALSE}
url <- paste0("https://raw.githubusercontent.com/cbrents/stat796logreg/master/hw2/icu.csv")
icu<-read_csv(url)
as.factor(icu$died)
```

## Question 1 
Consider the relationship between vital status (died, 1=died 0=alive) and systolic blood pressure (spb) at time of admission.

a. provide a plot that jointly summarizes these two variables.  Explain what trends in the data you can identify from this plot.

```{r echo=FALSE}
ggplot(icu) + 
  theme_bw() + 
  geom_line(aes(x = sbp, group = died, col = as.factor(died)),
            stat = "density") +
  ggtitle("Density estimates of systolic bloopd pressure by 
vital status (death or survival) during ICU stay") +
  labs(x="Systolic Blood Pressure (bpm)", y="Density", col="Vital Status During ICU Stay") +
  scale_color_hue(labels = c("0 = survival", "1 = death"))
```

Based on the density graph, we can see that patients who died in the ICU had spb between 100 and 160 bpm.  Survivors had a larger range of sbp - approximately 75 bpm to 150 bpm. 


b. fit a logistic regression model using vital status as the outcome and sbp as the predictor of interest. Provide the estimated regression coefficients. 

```{r, include=FALSE}
icu_glm <- glm(died~sbp, family=binomial, data=icu)
icu_glm
coef(icu_glm)
coef(icu_glm)[2] #sbp estimate
icucoef<- exp(coef(icu_glm))

#table(Estimated_Regression_Coefficients=icu_glm$coefficients) %>% 
 # kable
#kable(icu_glm$coefficients,
#      caption = "estimated regression coefficients")

```
The answer is `r icucoef`
A difference of one unit of SBP is associated with an odds ratio of exp(-0.777)= 1.02 for death during ICU stay. 

c. what is the estimated odds ratio for dying during an ICU visit, comparing those who differ in SBP at admission by 1mmHg?

```{r}

diff1mg<- exp(1*coef(icu_glm)[2]) #odds ratio for dying for 1 unit difference in SBP
```
The answer is `r diff1mg`

```{r practice, eval=FALSE}
x<- predict(icu_glm, newdata=data.frame(sbp=1), type="link")#link gives log odds 
exp(x) #exp log odds gives OR 
xx<- predict(icu_glm, newdata=data.frame(sbp=1), type="response")  #response gives predicted probability; link gives f(y) values
# https://www.statisticshowto.datasciencecentral.com/log-odds/
or<- xx/(1-xx)
kable(or, caption= "estimated OR for dying during an ICU visit, for SBP differing by 1mmHg")
```

d. What is the estimated odds ratio for dying during an ICU visit, comparing those who differ in SBP at admission by 15mmHg?

```{r}
diff15mg<- exp(15*coef(icu_glm)[2]) #odds ratio for dying for 15 mmHg difference in SBP
```
The answer is `r diff15mg`

```{r, eval=FALSE}
#newdata=something new
y<- predict(icu_glm, newdata=data.frame(sbp=15), type="response")  #response gives predicted probability; link gives f(y) values
yy<- y/(1-y)
exp(predict(icu_glm, newdata=data.frame(sbp=15), type="link"))  #same number doing long way! 
kable(yy, caption= "estimated OR for dying during an ICU visit, for SBP differing by 15mmHg")

```

e. Provide the code for 1bcd as an Appendix

##Question 2
Consider the relationship between vital status (died) and visit type (type)

a. Fill in a 2x2 table with the number of subjects having each combination of visit type and vital status

```{r}
#table(Died=icu$died, Type=icu$type)
```

 died   Elective  Emergency
---  ---------  ----------- ---
0    2               38     40
1    51             109     160
     53             147     200  
---  ---------  ----------- ---

```{r}
icu_proportion<- icu %>% 
  count(died, type) %>% 
  mutate(prop=prop.table(n))
#kable(icu_proportion)

b<- 51/(51+2)
c<- 109/(109+38)
# bc<-merge(b,c)
# colnames(bc)<-c("elective", "emergency")
# kable(bc,
#       caption="proportion of those who died whose visit was:")
```

b. Of those whose visit was elective, proportion of those who did.  The answer is `r b`
c. of those whose visit was emergency, proportion of those who died.  The anwer is `r c` 
d. fit a logistic regression model using vital status as outcome and visit type and the predictor of interest.

```{r}
icu2_glm<- glm(died~type, family=binomial, data=icu)
coefd<- coef(icu2_glm)
coef(icu2_glm)[2] #sbp estimate
exp(coef(icu2_glm))

#kable(icu2_glm$coefficients,     caption = "estimated regression coefficients")

```
Coefficients are `r coefd`

e. What is the estimated odds ratio for dying during an ICU visit, comparing those whose visit was an emergency to those whose was elective. 
```{r}
#1/(1+exp(-(coef(icu2_glm)[1] + coef(icu2_glm)[2]*0)))  #multiply by a number ; what about 0,1 variables? 
#exp(predict(icu2_glm, newdata=data.frame(died=1), type="link")) 
#remember what's on top, emergency =1

twoe<- exp(predict(icu2_glm, newdata=data.frame(type="emergency"), type="link")) #just looks at risk of having emergency surgey  #generates odds 
exp(coef(icu2_glm)[2])
#exp(predict(icu2_glm, newdata=data.frame(type="1"), type="link"))
```
The answer is `r twoe`

```{r}
f<- predict(icu2_glm, newdata = data.frame(type="elective"), type="response")
g<- predict(icu2_glm, newdata=data.frame(type="emergency"), type="response")
#fg<-merge(f,g)
#colnames(fg)<- c("elective", "emergency")
#kable(fg,
      #caption="predicted risk of dying during an ICU visit for people whose survey is:")
```
f. What is the predicted risk of dying during an ICU visit for people whose procedure is elective? The answer is `r f`
g. What is the predicted risk of dying during an ICU visit for people whose procedure is an emergency? The answer is `r g`

h. How do your answers for (f) and (g) compare to (b) and (c)? They're the same! 
    f. `r f`
    g. `r g`
    b. `r b`
    c. `r c`
    

```{r, eval=FALSE}
kable(bc,
      caption="proportion of those who died whose visit was:")
kable(fg,
      caption="predicted risk of dying during an ICU visit for people whose survey is:")


```
i. Provide the code you used for parts (d)-(g) as an Appendix after all questions.

##3. Again, consider the relationship between vital status and visit type.
a. create a new variable 'emergency' that takes the value of 1 when 'type' is 'emergency' and the value of 0 when 'type' is 'elective'. 
```{r}
icu_type<- icu %>%
  mutate(emergency = type,
         emergency=recode(emergency, "emergency"=1, "elective"=0)) 
icu_type$emergency<- as.numeric(icu_type$emergency)
#mutate , recode, old = new
```
 
b. Fit a logistic regression model using emergency as the outcome and vital status as the predictor of interest. Provide the coefficient estimates from the fit model.

```{r}
icu_type_glm <- glm(emergency~died, family=binomial, data=icu_type)
coefb<- coef(icu_type_glm)
exp(coef(icu_type_glm))

kable(icu_type_glm$coefficients,
      caption = "estimated regression coefficients")
```
Coefficients are `r coefb`

c. What is the estimated odds ratio for the procedure being emergency, comparing those who died during their ICU visit to those who did not.

emergency = 1 the outcome of interest  

```{r}
mypred <- exp(predict(icu_type_glm, newdata=data.frame(died=1), type="link"))  #response gives predicted probability; link gives f(y) values  #emergenc is outcome of interested, died = 1, variable of inerset 
#exp(coefficient of died =1 )
#mypred <- exp(predict(icu_type_glm, newdata=data.frame(type=1), type="link"))
#in line coding 

```
round command in line 

The value is `r mypred`

d. How does your response to 3(c) compare to part (e) of Question 2?

e. Provide the code you used for parts (a)-(c) as an Appendix after all questions.


## Appendix
```{r show-code, ref.label=all_labels(), echo = TRUE, eval=FALSE}
#should print all code
```



