---
title: "hw2"
author: "Colleen Brents"
date: "January 30, 2019"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
```
#question about 2e and 3d

```{r load libraries, include=FALSE}
library(ggplot2)
library(dplyr)
library(tidyr)
library(readr)
library(knitr)
library(kableExtra)
```

```{r data input, include=FALSE}
url <- paste0("https://raw.githubusercontent.com/cbrents/stat796logreg/master/hw2/icu.csv")
icu<-read_csv(url)
as.factor(icu$died)
```

## Question 1 
Consider the relationship between vital status (died, 1=died 0=alive) and systolic blood pressure (spb) at time of admission.

a. provide a plot that jointly summarizes these two variables.  Explain what trends in the data you can identify from this plot.

```{r echo=FALSE}
ggplot(icu) + 
  theme_bw() + 
  geom_line(aes(x = sbp, group = died, col = as.factor(died)),
            stat = "density") +
  ggtitle("Density estimates of systolic bloopd pressure by 
vital status (death or survival) during ICU stay") +
  labs(x="Systolic Blood Pressure (bpm)", y="Density", col="Vital Status During ICU Stay") +
  scale_color_hue(labels = c("0 = survival", "1 = death"))
```

Based on the density graph, we can see that patients who died in the ICU had spb between 100 and 160 bpm.  Survivors had a larger range of sbp - approximately 75 bpm to 150 bpm. 


b. fit a logistic regression model using vital status as the outcome and sbp as the predictor of interest. Provide the estimated regression coefficients. 

```{r}
icu_glm <- glm(died~sbp, family=binomial, data=icu)
icu_glm
coef(icu_glm)
coef(icu_glm)[2] #sbp estimate
exp(coef(icu_glm))

table(Estimated_Regression_Coefficients=icu_glm$coefficients) %>% 
  kable
kable(icu_glm$coefficients,
      caption = "estimated regression coefficients")

```
A difference of one unit of SBP is associated with an odds ratio of exp(-0.777)= 1.02 for death during ICU stay. 

c. what is the estimated odds ratio for dying during an ICU visit, comparing those who differ in SBP at admission by 1mmHg?

```{r}
x<- predict(icu_glm, newdata=data.frame(sbp=1), type="link")#link gives log odds 
exp(x) #exp log odds gives OR 
xx<- predict(icu_glm, newdata=data.frame(sbp=1), type="response")  #response gives predicted probability; link gives f(y) values
# https://www.statisticshowto.datasciencecentral.com/log-odds/
or<- xx/(1-xx)
kable(or, caption= "estimated OR for dying during an ICU visit, for SBP differing by 1mmHg")
```

d. What is the estimated odds ratio for dying during an ICU visit, comparing those who differ in SBP at admission by 15mmHg?

```{r}
y<- predict(icu_glm, newdata=data.frame(sbp=15), type="response")  #response gives predicted probability; link gives f(y) values
yy<- y/(1-y)
exp(predict(icu_glm, newdata=data.frame(sbp=15), type="link"))  #same number doing long way! 
kable(yy, caption= "estimated OR for dying during an ICU visit, for SBP differing by 15mmHg")

```

e. Provide the code for 1bcd as an Appendix

##Question 2
Consider the relationship between vital status (died) and visit type (type)

a. Fill in a 2x2 table with the number of subjects having each combination of visit type and vital status

```{r}
icu_2by2 <- table(Died=icu$died, Type=icu$type)
colnames(icu_2by2) <-c("Elective", "Emergency")
rownames(icu_2by2) <- c("Survived", "Died")
kable(icu_2by2)
```
b. Of those whose visit was elective, proportion of those who died 
c. of those whose visit was emergency, proportion of those who died 
```{r}
icu_proportion<- icu %>% 
  count(died, type) %>% 
  mutate(prop=prop.table(n))
kable(icu_proportion)

b<- 51/(51+2)
c<- 109/(109+38)
bc<-merge(b,c)
colnames(bc)<-c("elective", "emergency")
kable(bc,
      caption="proportion of those who died whose visit was:")
```

d. fit a logistic regression model using vital status as outcome and visit type and the predictor of interest.

```{r}
icu2_glm<- glm(died~type, family=binomial, data=icu)
coef(icu2_glm)
coef(icu2_glm)[2] #sbp estimate
exp(coef(icu2_glm))

kable(icu2_glm$coefficients,
      caption = "estimated regression coefficients")

```
e. What is the estimated odds ratio for dying during an ICU visit, comparing those whose visit was an emergency to those whose was elective.
```{r}
#1/(1+exp(-(coef(icu2_glm)[1] + coef(icu2_glm)[2]*0)))  #multiply by a number ; what about 0,1 variables? 
#exp(predict(icu2_glm, newdata=data.frame(died=1), type="link")) 

exp(predict(icu2_glm, newdata=data.frame(type="emergency"), type="link")) #just looks at risk of having emergency surgey

```


f. What is the predicted risk of dying during an ICU visit for people whose procedure is elective?
g. What is the predicted risk of dying during an ICU visit for people whose procedure is an emergency?
```{r}
f<- predict(icu2_glm, newdata = data.frame(type="elective"), type="response")
g<- predict(icu2_glm, newdata=data.frame(type="emergency"), type="response")
fg<-merge(f,g)
colnames(fg)<- c("elective", "emergency")
kable(fg,
      caption="predicted risk of dying during an ICU visit for people whose survey is:")
```

h. How do your answers for (f) and (g) compare to (b) and (c)? They're the same! 
```{r}
kable(bc,
      caption="proportion of those who died whose visit was:")
kable(fg,
      caption="predicted risk of dying during an ICU visit for people whose survey is:")


```
i. Provide the code you used for parts (d)-(g) as an Appendix after all questions.

##3. Again, consider the relationship between vital status and visit type.
a. create a new variable 'emergency' that takes the value of 1 when 'type' is 'emergency' and the value of 0 when 'type' is 'elective'. 
```{r}
icu_type<- icu %>%
  mutate(emergency = type,
         emergency=recode(emergency, "emergency"=1, "elective"=0)) 
icu_type$emergency<- as.numeric(icu_type$emergency)
#mutate , recode, old = new
```
 
b. Fit a logistic regression model using emergency as the outcome and vital status as the predictor of interest. Provide the coefficient estimates from the fit model.

```{r}
icu_type_glm <- glm(emergency~died, family=binomial, data=icu_type)
coef(icu_type_glm)
exp(coef(icu_type_glm))

kable(icu_type_glm$coefficients,
      caption = "estimated regression coefficients")
```

c. What is the estimated odds ratio for the procedure being emergency, comparing those who died during their ICU visit to those who did not.

emergency = 1 

```{r}
#exp(predict(icu_type_glm, newdata=data.frame(emergency=1), type="link"))  #response gives predicted probability; link gives f(y) values


```


d. How does your response to (c) compare to part (e) of Question 2?

e. Provide the code you used for parts (a)-(c) as an Appendix after all questions.


## Appendix
```{r show-code, ref.label=all_labels(), echo = TRUE, eval=FALSE}
#should print all code
```




